Towards an Ecological Trait-data Standard

### 2019-03-29 manuscript/manuscript.Rmd
fix layout and line breaks

<!-- 4. distributed trait data must be integrated --> 
Meanwhile, open and reproducible science has become mainstream: publication of research data without access restrictions, with structured metadata and in accordance to data standards to enable their reuse, has become the declared goal of an open biodiversity knowledge management (http://www.bouchoutdeclaration.org/) and is increasingly demanded by journals and public research funding agencies [ @allianceofgermanscienceorganisations10; @royalsocietysciencepolicycentre12; @allison15a]. <!--@emerson15; @swan12;>
As a result, an increasing number of individual research projects publish their primary data on general-purpose file hosting services, where no data standards are enforced upon the uploaded material [@wilkinson16]. 
It is thus likely that trait data will become increasingly available, but a lack of data and metadata standardisation will hamper the efficient reuse and synthesis of published datasets.

@@

Today, the online publication of such data is greatly facilitated by file hosting services (e.g., Figshare, Zenodo, Researchgate, Data Dryad), which warrant long-term accessibility, and citeability via DOIs, and govern data sharing via license statements.
These platforms offer the hosting of publicly accessible *file repositories* at low-cost or for free, which makes them attractive for small and intermediate-sized research projects that cannot dedicate extra resources for data management. 
<!--However, although open for download, the trait datasets on data repositories might be stored in proprietary (e.g., .xlsx, .docx) or binary (e.g., .pdf) data formats which make extraction tedious and dependent on specific commercial software, putting the long-term and open accessibility of these data at risk. -->

Most importantly, these platforms enable public hosting of data with very low quality-thresholds regarding *metadata* documentation and data standardisation. Thus, although open for download, the trait datasets on data repositories might be stored in variable tabular structures and labelled following self-defined terms which makes extraction and further use unnecessarily tedious. <!--As a result, published datasets greatly vary in structure, nomenclature and metadata quality. -->

>>>>

<!-- 4. distributed trait data must be integrated --> 
Meanwhile, open and reproducible science has become mainstream: publication of research data without access restrictions, with structured metadata and in accordance to data standards to enable their reuse, has become the declared goal of an open biodiversity knowledge management (http://www.bouchoutdeclaration.org/) and is increasingly demanded by journals and public research funding agencies [ @allianceofgermanscienceorganisations10; @royalsocietysciencepolicycentre12; @allison15a]. <!--@emerson15; @swan12;-->
As a result, an increasing number of individual research projects publish their primary data on general-purpose file hosting services, where no data standards are enforced upon the uploaded material [@wilkinson16]. 
It is thus likely that trait data will become increasingly available, but a lack of data and metadata standardisation will hamper the efficient reuse and synthesis of published datasets.

@@

Today, the online publication of such data is greatly facilitated by file hosting services (e.g., Figshare, Zenodo, Researchgate, Data Dryad), which warrant long-term accessibility, and citeability via DOIs, and govern data sharing via license statements.
These platforms offer the hosting of publicly accessible *file repositories* at low-cost or for free, which makes them attractive for small and intermediate-sized research projects that cannot dedicate extra resources for data management. 
<!--However, although open for download, the trait datasets on data repositories might be stored in proprietary (e.g., .xlsx, .docx) or binary (e.g., .pdf) data formats which make extraction tedious and dependent on specific commercial software, putting the long-term and open accessibility of these data at risk. -->
Most importantly, these platforms enable public hosting of data with very low quality-thresholds regarding *metadata* documentation and data standardisation. Thus, although open for download, the trait datasets on data repositories might be stored in variable tabular structures and labelled following self-defined terms which makes extraction and further use unnecessarily tedious. <!--As a result, published datasets greatly vary in structure, nomenclature and metadata quality. -->
